---
title: "homework-03"
author: "Mila Matich"
date: 2026-05-28
format: pdf
---

# Part 1. Homework Set Up

```{r}
#| label: loading-in-packages-and-data
#| message: FALSE


# loading in packages
library(tidyverse) # loading in tidyverse 
library(janitor) # loading in janitor 
library(readxl) # loading in readxl
library(ggeffects) # loading in ggeffects 

# reading in and storing data as object "kelp"
kelp <- read_csv("data/temp-kelp.csv") # data/ in order to direct the file path 
# reading in and storing personal data as object "personal"
personal <- read_csv("") # data/ in order to direct the file path 
```

# Part 2. Problems 

## Problem 1. Giant kelp fronds

### a. An appropriate test

In our study, we are trying to measure the strength of the relationship between two variables (ocean temperature and giant kelp frond elongation rate). In order to test the strength of this relationship, we should use a Pearson’s product-moment correlation and a Spearman’s rank correlation. Pearson's correlation test is the parametric version of the test, while Spearman's rank correlation is the non-parametric version that measures the correlation between ranks (rather than continuous variables).

### b. Create a visualization

```{r}
#| label: generating-linear-regression-model 
# fitting the linear model to generate the ribbon + linear line 
kelp_model <- lm(
  kelp_elong ~ temp_c, # formula: kelp growth rate as a function of temp
  data = kelp) # data from the kelp data set 

# generating predictor values for the linear model + ribbon 
kelp_preds <- ggpredict(
  kelp_model,        # model object
  terms = "temp_c")  # predictor (in quotation marks)

# generating the visualization
# base layer: ggplot 
ggplot(data = kelp, # data from kelp data set
       mapping = aes(x = temp_c, # x-axis = temp 
                     y = kelp_elong)) + # y-axis = growth rate 
  # first layer: geom_point 
  geom_point(size = 3,# changing point size 
             fill = "darkgreen", # adding data points in the color green
             shape = 21) + # point shape = circles 
  # second layer: ribbon representing confidence interval
  # using predictions data frame
  geom_ribbon(data = kelp_preds, # data from predicted data set 
              aes(x = x, # x = x in the data set
                  y = predicted, # y = y in the data set
                  ymin = conf.low, # min y point on ribbon set by CI
                  ymax = conf.high), # max y point on ribbon set by CI
                  alpha = 0.1, # opaqueness of the ribbon
                  fill = "green") + # changing ribbon color 
  # third layer: line representing model predictions
  # using predictions data frame
  geom_line(data = kelp_preds, # data from predicted data set 
            aes(x = x, # x = x in the data set
                y = predicted)) + # y = y in the data set
  # labeling x- and y-axes 
  labs(x = "Temperature (°C)", # x = "Temperature (°C)"
       y = "Giant Kelp Frond Elongation Rate (cm~day^{-1})") + # y = elongation rate
  # changing theme to be different than he default 
  theme_minimal()
```

### c. Check your assumptions and run your test

### d. Results communication

### e. Test implications

### f. f. Double check your own work.

## Problem 2. Personal data

### a. Updating your visualizations

### b. Captions

## Problem 3. Affective visualization

### a. Describe in words what an affective visualization could look like for your personal data (3-5 sentences).

### b. Create a sketch (on paper) of your idea.

### c. Make a draft of your visualization.

### d. Write an artist statement.

### e. Prep your materials to share in class.

## Problem 4. Statistical critique

### a. Revisit and summarize

### b. Visual clarity

### c. Aesthetic clarity

### d. Recommendations






